---
title: "SNSF Research Network"
date: "`r Sys.Date()`"
output:
  rmdformats::html_clean:
    highlight: tango
    code_folding: show
---

**Highlighting collaborations in SNSF grants running in 2017.**

```{r init, include=FALSE, cache=FALSE}
if (!require(pacman)) install.packages("pacman")
p_load("tidyverse", "data.table", "knitr", "rmdformats")
## Global options
options(max.print = "75")
knitr::opts_chunk$set(
  echo = TRUE,
	cache = TRUE,
  prompt = FALSE,
  tidy = FALSE,
  comment = NA,
  message = FALSE,
  warning = FALSE)
opts_knit$set(width = 75)
YEAR <- 2017
```

## Data

```{r data_download, eval=FALSE}
p3url <- c(
  "http://p3.snf.ch/P3Export/P3_GrantExport.csv",
  "http://p3.snf.ch/P3Export/P3_PersonExport.csv")
# "http://p3.snf.ch/P3Export/P3_CollaborationExport.csv")

purrr::walk2(p3url, saveas, download.file, quiet = TRUE)
```

You can download up-to-date SNSF grant data [here](http://p3.snf.ch/Pages/DataAndDocumentation.aspx).  Before reading the datasets in R, however, I recommend using MS Excel (or Libre/Open Office) to fix csv parsing problems.

```{r data_load}
grants <- fread("grants.csv")
people <- fread("people.csv")
```

`grants.csv` and `people.csv` correspond to `P3_GrantExport.csv` and `P3_PersonExport.csv`, respectively.  I made sure they read faultlessly and cleaned their headers with the `clean_names()` function from the [`janitor` package](https://github.com/sfirke/janitor).

## Grants connecting Switzerland

We want to highlight `r YEAR` grants connecting institutes from different parts of Switzerland and of the world.

```{r collaborations}
p_load("lubridate", "stringr")

# focus on a time range
trange <- ymd(c(str_c(YEAR, "-01-01"), str_c(YEAR, "-12-31")))

# find grants running in that time range
grants <- grants %>%
  select(
    project_number, discipline_number, start_date, end_date) %>%
  mutate(
    start_date = ymd(str_sub(start_date, 1, 10)),
    end_date = ymd(str_sub(end_date, 1, 10)),
    domain = as.integer(str_extract_all(discipline_number, "^[0-9]"))) %>%
  filter(start_date <= trange[2], end_date >= trange[1])

head(grants)
# find the people involved in those grants
people <- people %>%
  select(person_id_snsf, institute_place, starts_with("projects")) %>%
  filter(institute_place != "") %>% 
  select(-projects_as_responsible_applicant) %>%
  unite(project_number, starts_with("projects"), sep = ";") %>%
  mutate(
    institute_place = str_replace_all(institute_place, " Cedex(.*)?| [0-9]{1,2}", ""),
    project_number = str_replace_all(project_number, "[;]+", ";"),
    project_number = str_replace_all(project_number, "^[;]|[;]$|NA", ""),
    project_number = str_split(project_number, ";")) %>% 
  unnest() %>%
  mutate(project_number = as.integer(project_number)) %>%
  semi_join(grants, by = "project_number")

head(people)
```


## The collaboration network

To show the network in a map, we focus on grants connecting people in different places.

```{r core}
# how many distinct places per grant?
places_per_grant <- group_by(people, project_number) %>%
  summarise(n_places = n_distinct(institute_place))

head(places_per_grant)
# network core: projects involving multiple places (n_places > 1)
core <- people %>%
  select(-person_id_snsf) %>%
  semi_join(filter(places_per_grant, n_places > 1), by = "project_number") %>%
  arrange(project_number) %>%
  distinct()

head(core)
```

We also need the geocodes of all the places of the network.  You find more information on how to get geocode data in `R` [here](https://gis.stackexchange.com/questions/158328/batch-geocoding-in-r) and [here](https://github.com/hrbrmstr/nominatim).  In essence, I use two APIs: `openstreetmaps` and `googe maps` as fallback.  To save the effort of writing html requests, I use dedicated `R` packages to query map data. To speed up the analysis, we will store the geocodes in `geocodes.csv`.

```{r geocodes, eval=FALSE}
# initialize geocodes as a tibble
geocodes <- tibble(
  place = sort(unique(core$institute_place)),
  lat = rep(NA_real_, n_distinct(core$institute_place)),
  lon = rep(NA_real_, n_distinct(core$institute_place)),
  addr = rep(NA_character_, n_distinct(core$institute_place)),
  id = rep(NA_character_, n_distinct(core$institute_place)))

p_load_gh("hrbrmstr/nominatim", "ggmap") # ggmap requires `libpng16-dev` in ubuntu
osm_key <- readLines("mapquest.key")

# openstreetmap api
osm <- function(query, osm_key) {
  r <- osm_search_spatial(query, limit = 1, key = osm_key)
  if (!is.null(r[[1]])) {
    c(r[[1]]$place_id, r[[1]]$display_name, r[[1]]$lat, r[[1]]$lon)
  } else return(NA)
}

# googlemaps api (2500 reqs/day, 50 reqs/sec max)
google <- function(query) {
  r <- geocode(query, output = "all")
  if (r$status == "OK") {
    c(r$results[[1]]$place_id, r$results[[1]]$formatted_address,
      r$results[[1]]$geometry$location$lat, r$results[[1]]$geometry$location$lng)
  } else return(NA)
}

for (k in seq_along(geocodes$place)) {
  cat("   ........ ", k, ": ", geocodes$place[k], "\n")
  info <- osm(geocodes$place[k], osm_key)
  cat("   osm info: ", info, "\n")
  # fallback
  if (is.na(info)) {
    info <- google(geocodes$place[k])
    cat("    google: ", info, "\n")
  }
  # store info in meta
  if (!is.na(info)) {
    geocodes$id[k] = info[1]
    geocodes$addr[k] = info[2]
    geocodes$lat[k] = as.numeric(info[3])
    geocodes$lon[k] = as.numeric(info[4])
  }
}
# it works despite of some warnings...
write_csv(geocodes, path = "geocodes.csv")
```

```{r metadata}
geocodes <- fread("geocodes.csv") %>% drop_na()

head(geocodes)
# add geocode data to the network core
core <- left_join(core, geocodes, by = c("institute_place" = "place"))
```

### Swiss collaboration network

Lets show the Swiss network first.  
The visualization is heavily inspired from [here](http://kateto.net/network-visualization).

```{r swisscore}
core_ch <- filter(core, str_detect(addr, "Switzerland$"))
# find projects on more than one place in Switzerland
ch_per_grant <- group_by(core_ch, project_number) %>%
  summarise(n_places = n_distinct(institute_place))
# swiss core: projects involving multiple places in Switzerland
core_ch <- semi_join(core_ch, filter(ch_per_grant, n_places > 1), by = "project_number")

# nodes as places
nodes <- group_by(core_ch, institute_place) %>%
  summarise(y = head(lat, 1), x = head(lon, 1), size = n()) %>%
  arrange(desc(size))

head(nodes)
# edges as grants
p_load("magrittr")
project_edges <- function(number, df) {
  edges <- filter(df, project_number == number) %>%
    select(institute_place) %$%
    combn(sort(institute_place), m = 2)
  tibble(from = edges[1, ],
         to = edges[2,]) %>%
    mutate(number = number) %>%
    select(number, everything())
}

# TODO: speed-up project_edges()
edges <- purrr::map(unique(core_ch$project_number), project_edges, df = core_ch)
edges <- do.call(bind_rows, edges)

# add geocode metadata and group
edges <- edges %>%
  left_join(select(nodes, -size), by = c("from" = "institute_place")) %>%
  rename(x1 = x, y1 = y) %>%
  left_join(select(nodes, -size), by = c("to" = "institute_place")) %>%
  rename(x2 = x, y2 = y) %>%
  group_by(from, to) %>%
  summarise(
    x1 = head(x1, 1),
    y1 = head(y1, 1),
    x2 = head(x2, 1),
    y2 = head(y2, 1),
    strength = n())

head(edges)
# plot the nodes and the edges on the map of Switzerland
p_load("maps", "mapdata", "geosphere")

col_dark <- adjustcolor("#800026", alpha = 0.3)
col_light <- adjustcolor("#feb24c", alpha = 0.3)
edge_pal <- colorRampPalette(c(col_dark, col_light), alpha = TRUE)
edge_col <- edge_pal(100)

svg("core_ch.svg")
par(mar = c(0, 0, 0, 0))
maps::map(database = "worldHires", regions = "Switzerland",
          fill = FALSE, col = rgb(0, 0, 0, .2))

# map nodes
points(x = nodes$x, y = nodes$y, pch = 16,
  cex = log(nodes$size)/2, col = rgb(0, 0, 0, .2))
# map edges
for (k in 1:nrow(edges))  {
  arc <- gcIntermediate(
    c(edges$x1[k], edges$y1[k]),
    c(edges$x2[k], edges$y2[k]),
    n = 100, addStartEnd = TRUE)
  strength <- round(100 * edges$strength[k] / max(edges$strength))
  lines(arc, col = edge_col[strength], lwd = sqrt(strength) / 2)
}
# label main cities
cities <- c("Zürich", "Lausanne", "Bern", "Genève", "Basel", "Fribourg", "Neuchâtel", "St. Gallen", "Lugano", "Luzern", "Winterthur")

city_nodes <- filter(nodes, institute_place %in% cities)
graphics::text(
  city_nodes$x, city_nodes$y, 
  labels = city_nodes$institute_place, pos = 3, cex = .7, col = rgb(0, 0, 0, .75))
dev.off()
```

![switzerland](core_ch.svg)

## About 

This page uses the `html_cleam` template from the [`prettydoc` package](https://github.com/yixuan/prettydoc).
